Variants of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus

### 2016-08-14 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Revier 1. (1) Koon changed to Koons
Reviewer 1: (2) G(x) changed to G(\gamma)
Reviewer 2:  (D.7) Section 4, Proof of Theorem 2
(D.8) Section 4, Proof of Lemma 1
(D.9) Section 4, Proof of Lemma 2
(D.10) Section 5, Proof of Lemma 1
(D.11) Section 6, Proof of Theorem 1
(D.4) Section 3

From an axiomatic point of view, modal logics differ with respect to which additional axiom schemas they admit. Some common axiom schemas of relevance to the ontological argument are T ($\nec A \imp A$), B ($A \imp \nec \pos A$), 4 ($\nec A \imp \nec \nec A $) and 5 ($\pos A \imp \nec \pos A$). From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accessibility relation in Kripke models \citep{negri}: T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity. In the models of logic {\bf S5}, all these axioms are satisfied; in the models of logic {\bf KB}, K and B are satisfied; and in the models of logic {\bf B}, K, T and B are satisfied.

@@

	\AXC{$ $} \RightLabel{5}
	\UIC{$\rho(\gamma)$}

		\AXC{$ $} \RightLabel{2}
		\UIC{$\neg P(\rho)$} 
				\AXC{Axiom 1} \dashedLine
        \UIC{$\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]$} \RightLabel{$\all_E$}
        \UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
        \UIC{$\neg P(\rho) \imp P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
			\BIC{$P(\lambda x. \neg \rho(x))$}

					\AXC{$ $} \RightLabel{1}
					\UIC{$G(\gamma)$} \dottedLine
					\UIC{$\all \varphi. P(\varphi) \imp \varphi(\gamma)$} \RightLabel{$\all_E$}
					\UIC{$P(\lambda x. \neg \rho(x)) \imp \neg \rho(\gamma)$} \RightLabel{$\imp_E$}
				\BIC{$\neg \rho(\gamma)$} \RightLabel{$\neg_E$}      
	   \BIC{$\bot$} \RightLabel{$\neg_I^2$}
	   \UIC{$\neg\neg P(\rho)$} \RightLabel{$\neg\neg_E$}
	   \UIC{$[\subproof_3] \qquad P(\rho)$}
	   \end{prooftree}
\end{scriptsize}

\begin{scriptsize}
\begin{prooftree}
\AXC{$ $} \RightLabel{1}
\UIC{$G(\gamma)$}

 \AXC{$\subproof_3$}\dashedLine
 \UIC{$P(\rho)$}
	      \AXC{Axiom 4} \dashedLine
        \UIC{$\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]$} \RightLabel{$\all_E$}
	   	  \UIC{$P(\rho) \imp \nec P(\rho)$} \RightLabel{$\imp_E$}
		 \BIC{$\nec P(\rho)\scott$}
			\AXC{$ $} \RightLabel{4}
			\UIC{$\omega: P(\rho)$}
					\AXC{$ $} \RightLabel{3}
					\UIC{$\omega: G(\delta) $} \dottedLine
					\UIC{$\omega: \all \varphi. P(\varphi) \imp \varphi(\delta)$} \RightLabel{$\all_E$} 
					\UIC{$\omega: P(\rho) \imp \rho(\delta)$} \RightLabel{$\imp_E$}
				\BIC{$\omega: \rho(\delta) $} \RightLabel{$\imp_I^3$}
				\UIC{$\omega: G(\delta) \imp \rho(\delta) $} \RightLabel{$\all_I$}
				\UIC{$\omega: \all y. G(y) \imp \rho(y) $} \RightLabel{$\imp_I^4$}
				\UIC{$\omega: P(\rho) \imp (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\nec_I$}
				\UIC{$\nec (P(\rho) \imp (\all y. G(y) \imp \rho(y) ))$} \RightLabel{$\imp_E$}
						\AXC{ K} \RightLabel{$\imp_E$}
					\BIC{$\nec P(\rho) \imp \nec (\all y. G(y) \imp \rho(y) )$}
			\BIC{$\nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_I^5$}
			\UIC{$\rho(\gamma) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\all_I$}
			\UIC{$\all \varphi. \varphi(\gamma) \imp \nec (\all y. G(y) \imp \varphi(y) )$} \RightLabel{$\wedge_I$}
		\BIC{$ G(\gamma) \wedge (\all \varphi. \varphi(\gamma) \imp \nec (\all y. G(y) \imp \varphi(y) ))$} \dottedLine
		\UIC{$\ess{G}{\gamma}$} \RightLabel{$ \imp_I^1$}
		\UIC{$G(\gamma) \imp \ess{G}{\gamma}$} \RightLabel{$\all_I$}
		\UIC{$\all x. G(x) \imp \ess{G}{x}$}
\end{prooftree}
\end{scriptsize}

@@

\begin{prooftree}
\AXC{$ $} \RightLabel{1}
\UIC{$\ex z. G(z)$} \RightLabel{$\ex_E$}
\UIC{$[\subproof_4] \qquad G(\gamma)$}
\end{prooftree}

\begin{prooftree}
\AXC{$\subproof_4$} \dashedLine
\UIC{$G(\gamma)$}
		\AXC{Theorem 2} \dashedLine 
		\UIC{$\all x. G(x) \imp \ess{G}{x}$} \RightLabel{$\all_E$}
		\UIC{$G(\gamma) \imp \ess{G}{\gamma}$}
	\BIC{$\ess{G}{\gamma}\scott$}
				\AXC{Axiom 5} \dashedLine
				\UIC{$P(E)$}
						\AXC{$\subproof_4$} \dashedLine
						\UIC{$G(\gamma)$} \dottedLine
						\UIC{$\all \varphi. P(\varphi) \imp \varphi(\gamma)$} \RightLabel{$\all_E$}
						\UIC{$P(E) \imp E(\gamma)$} 
					\BIC{$E(\gamma)\scott$} \dottedLine
					\UIC{$\all \varphi. \ess{\varphi}{\gamma} \imp \nec \ex x. \varphi(x)$} \RightLabel{$\all_E$}
					\UIC{$\ess{G}{\gamma} \imp \nec \ex x. G(x)$}
		\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
		\UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}

@@

\AXC{$ $} \RightLabel{1}
\UIC{$\pos \ex x. G(x)$} 
		\AXC{Lemma 1} \dashedLine
		\UIC{$\omega: \ex x. G(x) \imp \pos \nec \ex x. G(x)$} \RightLabel{$\nec_I$}
		\UIC{$\nec (\ex x. G(x) \imp \nec \ex x. G(x)) $}
				\AXC{Distribution Principle}
			\BIC{$\pos \ex x. G(x) \imp \pos \nec \ex x. G(x)$} \RightLabel{$\imp_E$}
	\BIC{$\pos \nec \ex x. G(x)\scott$}
			\AXC{S5 Iteration Principle} \dashedLine
			\UIC{$\pos \nec \ex x. G(x) \imp \nec \ex x. G(x)$}
		\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
		\UIC{$\pos \ex z. G(z) \imp \nec \ex x. G(x)$}
\end{prooftree}
\end{scriptsize}

@@

\begin{proof} \hfill
\begin{prooftree}
\AXC{$ $} \RightLabel{1}
\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
\UIC{$[\subproof_6] \qquad G(\gamma)$}
\end{prooftree}

\begin{tiny}
\begin{prooftree}
\AXC{$ $} \RightLabel{2}
\UIC{$\neg P(\rho)$} 
\AXC{Axiom 1} \dashedLine
\UIC{$\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]$} \RightLabel{$\all_E$}
\UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
\UIC{$\neg P(\rho) \imp P(\lambda x. \neg \rho(x))$}\RightLabel{$\imp_E$}
\BIC{$P(\lambda x. \neg \rho(x))$}
\AXC{$\subproof_6$} \dashedLine
\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{N:D1}}
\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(\gamma))$}\RightLabel{$\forall_E$}
\UIC{$P(\lambda x. \neg \rho(x)) \imp \neg \rho(\gamma)$}\RightLabel{$\imp_E$}
\BIC{$\neg \rho(\gamma)$}
\AXC{$ $} \RightLabel{3}
\UIC{$\rho(\gamma)$} \RightLabel{$\imp_E$}
\BIC{$\bot$}\RightLabel{$\imp_I^2$}
\UIC{$\neg\neg P(\rho)$}\RightLabel{$\neg\neg_E$}
\UIC{$P(\rho)$}
\AXC{Axiom \ref{N:A4}} \dashedLine
\UIC{$\all \varphi.[P(\varphi) \imp \nec\; P(\varphi)]$}\RightLabel{$\forall_E$}
\UIC{$P(\rho) \imp \nec\; P(\rho)$}\RightLabel{$\imp_E$}
\BIC{$\nec P(\rho)$}\RightLabel{$\imp_I^3$}
\UIC{$[\subproof_7] \qquad \rho(\gamma) \imp \nec\; P(\rho)$}
\end{prooftree}
\end{tiny}

\begin{small}
\begin{prooftree}
\AXC{$ $} \RightLabel{5}
\UIC{$\nec P(\rho)$} \RightLabel{$\nec_E$}
\UIC{$\omega: P(\rho)$}
\AXC{$ $} \RightLabel{4}
\UIC{$\omega: G(y)$}  \dottedLine \RightLabel{D\ref{N:D1}}
\UIC{$\omega: \forall \varphi.(P(\varphi) \imp \varphi(y))$}\RightLabel{$\forall_E$}
\UIC{$\omega: P(\rho) \imp \rho(y)$}\RightLabel{$\imp_E$}
\BIC{$\omega: \rho(y)$} \RightLabel{$\imp_I^4$}
\UIC{$\omega: G(y) \imp \rho(y)$}\RightLabel{$\forall_I$}
\UIC{$\omega: \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\nec_I$}
\UIC{$\nec \forall y.(G(y) \imp \rho(y))$} \RightLabel{$\imp_I^5$}
\UIC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$} \RightLabel{$\imp_E$}
\AXC{$ $} \RightLabel{6}
\UIC{$\rho(\gamma)$} \RightLabel{}
\AXC{$\subproof_7$} \dashedLine
\UIC{$\rho(\gamma)\imp\Box P(\rho)$}\RightLabel{$\imp_E$}
\BIC{$\nec P(\rho)$} \RightLabel{$\imp_E$}
\BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^6$}
\UIC{$\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
\UIC{$\forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}
\AXC{$\subproof_6$} \dashedLine
\UIC{$G(\gamma)$}  
\BIC{$G(\gamma) \wedge \forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{N:D2}}
\UIC{$[\subproof_8] \qquad \ess{G}{\gamma}$}
\end{prooftree}
\end{small}

\begin{small}
\begin{prooftree}
\AXC{$\subproof_8$} \dashedLine
\UIC{$\ess{G}{\gamma}$}
\AXC{Axiom \ref{A5}} \dashedLine
\UIC{$P(E)$}\dottedLine \RightLabel{D\ref{N:D3}}
\AXC{$\subproof_6$} \dashedLine
\UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{N:D1}}
\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(\gamma))$}\RightLabel{$\forall_E$}
\UIC{$P(E) \imp E(\gamma)$}\RightLabel{$\imp_E$}
\BIC{$E(\gamma)$}\dottedLine \RightLabel{D\ref{N:D3}}
\UIC{$ \all \varphi.[\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
\UIC{$ \ess{G}{\gamma} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
\UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
\end{prooftree}
\end{small}
\end{proof}

@@

\begin{small}
\begin{prooftree}
\AXC{Theorem 2}\dashedLine
\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\doubleLine
\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$}
\UIC{$\all y.[G(y) \imp \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$} \doubleLine
%\UIC{$\all y.[G(y) \imp (A(y) \imp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
\UIC{$\all y.[G(y) \imp (A \imp \nec \all x. (G(x) \imp A))]$} \doubleLine
\UIC{$[\subproof_9] \qquad \ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
\end{prooftree}
\end{small}

@@

In G\"odel's ontological proof, we are proving a restricted modal collapse, which applies to one specific formula, the existence of a God-like being. The interest in the proof naturally decreases if a consequence of the axiomatization is a modal collapse for \emph{all} formulas. Therefore, an improvement would be obtained if the modal collapse was limited only to the property of being god-like or at least to a restricted collection of properties. Several solutions to the problem of modal collapse have been proposed.
Fitting, for instance, has argued that greater care has to be taken with the semantics of higher-order modal logics. Quantified variables may be rigid or flexible; and properties may be treated as intensional or extensional. Making the right choices may prevent the modal collapse \citep{fitting}[Sections 11.9 and 11.10]. Koon \citep{koons}, Anderson \citep{and90}[p. 292] and Sobel \citep{sobel2}[p. 133] also discuss the idea that the notion of property over which quantification is allowed might be too general and restrictions might be appropriate. Two alternative and more conservative solutions, requiring only modifications of the axioms and definitions but no change in the logic, are discussed in greater detail in the next sections.

@@

\begin{prooftree}
\AXC{$\subproof_6$} \dashedLine
\UIC{$G(\gamma)$} \RightLabel{}\dottedLine \RightLabel{D\ref{A:D1}}
\UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
\UIC{$P(\lambda x.\rho(x)) \biimp \nec \rho(\gamma)$}
\AXC{$ $} \RightLabel{6}
\UIC{$\nec\rho(\gamma)$} \RightLabel{}
\BIC{$P(\rho)$} \RightLabel{$\imp_E$}

@@

\begin{prooftree}
\AXC{$ $} \RightLabel{5}
\UIC{$\nec P(\rho)$} \RightLabel{$\nec_E$}
\UIC{$\omega: P(\rho)$}
\AXC{$ $} \RightLabel{4}
\UIC{$\omega: G(y)$}  \dottedLine \RightLabel{D\ref{A:D1}}
\UIC{$\omega: \forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
\UIC{$\omega: P(\rho) \biimp \nec \rho(y)$}\RightLabel{$\biimp_E$}
\BIC{$\omega: \nec \rho(y)$}\RightLabel{$\nec_E$}
\UIC{$\omega: \rho(y)$} \RightLabel{$\imp_I^4$}
\UIC{$\omega: G(y) \imp \rho(y)$}\RightLabel{$\forall_I$}
\UIC{$\omega: \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\nec_I$}
\UIC{$\nec \forall y.(G(y) \imp \rho(y))$} \RightLabel{$\imp_I^5$}
\UIC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$} \RightLabel{$\imp_E$}

\AXC{$\subproof_7$} \dashedLine
\UIC{$\nec P(\rho)$}
\BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^6$}
\UIC{$[\subproof_8] \qquad \nec\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
\end{prooftree}

>>>>

From an axiomatic point of view, modal logics differ with respect to which additional axiom schemas they admit. Some common axiom schemas of relevance to the ontological argument are T ($\nec A \imp A$), B ($A \imp \nec \pos A$), 4 ($\nec A \imp \nec \nec A $) and 5 ($\pos A \imp \nec \pos A$). From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accessibility relation in Kripke models (See for example \citep{fitting2}): T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity. In the models of logic {\bf S5}, all these axioms are satisfied; in the models of logic {\bf KB}, K and B are satisfied; and in the models of logic {\bf B}, K, T and B are satisfied.

@@

	\AXC{$ $} \RightLabel{4}
	\UIC{$\rho(\gamma)$}

		\AXC{$ $} \RightLabel{1}
		\UIC{$\neg P(\rho)$} 
				\AXC{Axiom 1} \dashedLine
        \UIC{$\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]$} \RightLabel{$\all_E$}
        \UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
        \UIC{$\neg P(\rho) \imp P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
			\BIC{$P(\lambda x. \neg \rho(x))$}

					\AXC{$ $} \RightLabel{5}
					\UIC{$G(\gamma)$} \dottedLine
					\UIC{$\all \varphi. P(\varphi) \imp \varphi(\gamma)$} \RightLabel{$\all_E$}
					\UIC{$P(\lambda x. \neg \rho(x)) \imp \neg \rho(\gamma)$} \RightLabel{$\imp_E$}
				\BIC{$\neg \rho(\gamma)$} \RightLabel{$\neg_E$}      
	   \BIC{$\bot$} \RightLabel{$\neg_I^1$}
	   \UIC{$\neg\neg P(\rho)$} \RightLabel{$\neg\neg_E$}
	   \UIC{$[\subproof_3] \qquad P(\rho)$}
	   \end{prooftree}
\end{scriptsize}

\begin{scriptsize}
\begin{prooftree}
\AXC{$ $} \RightLabel{5}
\UIC{$G(\gamma)$}

 \AXC{$\subproof_3$}\dashedLine
 \UIC{$P(\rho)$}
	      \AXC{Axiom 4} \dashedLine
        \UIC{$\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]$} \RightLabel{$\all_E$}
	   	  \UIC{$P(\rho) \imp \nec P(\rho)$} \RightLabel{$\imp_E$}
		 \BIC{$\nec P(\rho)\scott$}
			\AXC{$ $} \RightLabel{3}
			\UIC{$\omega: P(\rho)$}
					\AXC{$ $} \RightLabel{2}
					\UIC{$\omega: G(\delta) $} \dottedLine
					\UIC{$\omega: \all \varphi. P(\varphi) \imp \varphi(\delta)$} \RightLabel{$\all_E$} 
					\UIC{$\omega: P(\rho) \imp \rho(\delta)$} \RightLabel{$\imp_E$}
				\BIC{$\omega: \rho(\delta) $} \RightLabel{$\imp_I^2$}
				\UIC{$\omega: G(\delta) \imp \rho(\delta) $} \RightLabel{$\all_I$}
				\UIC{$\omega: \all y. G(y) \imp \rho(y) $} \RightLabel{$\imp_I^3$}
				\UIC{$\omega: P(\rho) \imp (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\nec_I$}
				\UIC{$\nec (P(\rho) \imp (\all y. G(y) \imp \rho(y) ))$} \RightLabel{$\imp_E$}
						\AXC{Axiom K} \RightLabel{$\imp_E$}
					\BIC{$\nec P(\rho) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_E$}
			\BIC{$\nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_I^4$}
			\UIC{$\rho(\gamma) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\all_I$}
			\UIC{$\all \varphi. \varphi(\gamma) \imp \nec (\all y. G(y) \imp \varphi(y) )$} \RightLabel{$\wedge_I$}
		\BIC{$ G(\gamma) \wedge (\all \varphi. \varphi(\gamma) \imp \nec (\all y. G(y) \imp \varphi(y) ))$} \dottedLine
		\UIC{$\ess{G}{\gamma}$} \RightLabel{$ \imp_I^5$}
		\UIC{$G(\gamma) \imp \ess{G}{\gamma}$} \RightLabel{$\all_I$}
		\UIC{$\all x. G(x) \imp \ess{G}{x}$}
\end{prooftree}
\end{scriptsize}

@@

\begin{prooftree}
\AXC{$ $} \RightLabel{1}
\UIC{$\ex z. G(z)$} \RightLabel{$\ex_E$}
\UIC{$G(\gamma)$}
		\AXC{Theorem 2} \dashedLine 
		\UIC{$\all x. G(x) \imp \ess{G}{x}$} \RightLabel{$\all_E$}
		\UIC{$G(\gamma) \imp \ess{G}{\gamma}$}  \RightLabel{$\imp_E$}
	\BIC{$\ess{G}{\gamma}\scott$}
				\AXC{Axiom 5} \dashedLine
				\UIC{$P(E)$}
						\AXC{$ $} \RightLabel{1}
						\UIC{$\ex z. G(z)$} \RightLabel{$\ex_E$}
						\UIC{$G(\gamma)$} \dottedLine
						\UIC{$\all \varphi. P(\varphi) \imp \varphi(\gamma)$} \RightLabel{$\all_E$}
						\UIC{$P(E) \imp E(\gamma)$} \RightLabel{$\imp_E$}
					\BIC{$E(\gamma)\scott$} \dottedLine
					\UIC{$\all \varphi. \ess{\varphi}{\gamma} \imp \nec \ex x. \varphi(x)$} \RightLabel{$\all_E$}
					\UIC{$\ess{G}{\gamma} \imp \nec \ex x. G(x)$} \RightLabel{$\imp_E$}
		\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
		\UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}

@@

\AXC{$ $} \RightLabel{1}
\UIC{$\pos \ex x. G(x)$} 
		\AXC{Lemma 1} \dashedLine
		\UIC{$\omega: \ex x. G(x) \imp \nec \ex x. G(x)$} \RightLabel{$\nec_I$}
		\UIC{$\nec (\ex x. G(x) \imp \nec \ex x. G(x)) $}
				\AXC{Distribution Principle}
			\BIC{$\pos \ex x. G(x) \imp \pos \nec \ex x. G(x)$} \RightLabel{$\imp_E$}
	\BIC{$\pos \nec \ex x. G(x)\scott$}
			\AXC{S5 Iteration Principle} \dashedLine
			\UIC{$\pos \nec \ex x. G(x) \imp \nec \ex x. G(x)$}
		\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
		\UIC{$\pos \ex z. G(z) \imp \nec \ex x. G(x)$}
\end{prooftree}
\end{scriptsize}

@@

\begin{proof} \hfill

\begin{small}
\begin{prooftree}
\AXC{$ $} \RightLabel{2}
\UIC{$\neg P(\rho)$} 
\AXC{Axiom 1} \dashedLine
\UIC{$\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]$} \RightLabel{$\all_E$}
\UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
\UIC{$\neg P(\rho) \imp P(\lambda x. \neg \rho(x))$}\RightLabel{$\imp_E$}
\BIC{$P(\lambda x. \neg \rho(x))$}
\AXC{$ $} \RightLabel{1}
\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{N:D1}}
\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(\gamma))$}\RightLabel{$\forall_E$}
\UIC{$P(\lambda x. \neg \rho(x)) \imp \neg \rho(\gamma)$}\RightLabel{$\imp_E$}
\BIC{$\neg \rho(\gamma)$}
\AXC{$ $} \RightLabel{3}
\UIC{$\rho(\gamma)$} \RightLabel{$\imp_E$}
\BIC{$\bot$}\RightLabel{$\imp_I^2$}
\UIC{$\neg\neg P(\rho)$}\RightLabel{$\neg\neg_E$}
\UIC{$P(\rho)$}
\AXC{Axiom \ref{N:A4}} \dashedLine
\UIC{$\all \varphi.[P(\varphi) \imp \nec\; P(\varphi)]$}\RightLabel{$\forall_E$}
\UIC{$P(\rho) \imp \nec\; P(\rho)$}\RightLabel{$\imp_E$}
\BIC{$\nec P(\rho)$}\RightLabel{$\imp_I^3$}
\UIC{$[\subproof_7] \qquad \rho(\gamma) \imp \nec\; P(\rho)$}
\end{prooftree}
\end{small}

\begin{small}
\begin{prooftree}
\AXC{$ $} \RightLabel{5}
\UIC{$\nec P(\rho)$} \RightLabel{$\nec_E$}
\UIC{$\omega: P(\rho)$}
\AXC{$ $} \RightLabel{4}
\UIC{$\omega: G(y)$}  \dottedLine \RightLabel{D\ref{N:D1}}
\UIC{$\omega: \forall \varphi.(P(\varphi) \imp \varphi(y))$}\RightLabel{$\forall_E$}
\UIC{$\omega: P(\rho) \imp \rho(y)$}\RightLabel{$\imp_E$}
\BIC{$\omega: \rho(y)$} \RightLabel{$\imp_I^4$}
\UIC{$\omega: G(y) \imp \rho(y)$}\RightLabel{$\forall_I$}
\UIC{$\omega: \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\nec_I$}
\UIC{$\nec \forall y.(G(y) \imp \rho(y))$} \RightLabel{$\imp_I^5$}
\UIC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$} \RightLabel{$\imp_E$}
\AXC{$ $} \RightLabel{6}
\UIC{$\rho(\gamma)$} \RightLabel{}
\AXC{$\subproof_7$} \dashedLine
\UIC{$\rho(\gamma)\imp\Box P(\rho)$}\RightLabel{$\imp_E$}
\BIC{$\nec P(\rho)$} \RightLabel{$\imp_E$}
\BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^6$}
\UIC{$\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
\UIC{$\forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}
\AXC{$ $} \RightLabel{1}
\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
\UIC{$G(\gamma)$}  \RightLabel{$\wedge_I$}
\BIC{$G(\gamma) \wedge \forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{N:D2}}
\UIC{$[\subproof_8] \qquad \ess{G}{\gamma}$}
\end{prooftree}
\end{small}

\begin{small}
\begin{prooftree}
\AXC{$\subproof_8$} \dashedLine
\UIC{$\ess{G}{\gamma}$}
\AXC{Axiom \ref{A5}} \dashedLine
\UIC{$P(E)$}\dottedLine 
\AXC{$ $} \RightLabel{1}
\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
\UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{N:D1}}
\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(\gamma))$}\RightLabel{$\forall_E$}
\UIC{$P(E) \imp E(\gamma)$}\RightLabel{$\imp_E$}
\BIC{$E(\gamma)$}\dottedLine \RightLabel{D\ref{N:D3}}
\UIC{$ \all \varphi.[\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
\UIC{$ \ess{G}{\gamma} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
\UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
\end{prooftree}
\end{small}
\end{proof}

@@

\begin{small}
\begin{prooftree}
\AXC{Theorem 2}\dashedLine
\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\dottedLine
\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$}\doubleLine
\UIC{$\all y.[G(y) \imp \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$} \doubleLine
%\UIC{$\all y.[G(y) \imp (A(y) \imp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
\UIC{$\all y.[G(y) \imp (A \imp \nec \all x. (G(x) \imp A))]$} \doubleLine
\UIC{$[\subproof_9] \qquad \ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
\end{prooftree}
\end{small}

@@

In G\"odel's ontological proof, we are proving a restricted modal collapse, which applies to one specific formula, the existence of a God-like being. The interest in the proof naturally decreases if a consequence of the axiomatization is a modal collapse for \emph{all} formulas. Therefore, an improvement would be obtained if the modal collapse was limited only to the property of being god-like or at least to a restricted collection of properties. Several solutions to the problem of modal collapse have been proposed.
Fitting, for instance, has argued that greater care has to be taken with the semantics of higher-order modal logics. Quantified variables may be rigid or flexible; and properties may be treated as intensional or extensional. Making the right choices may prevent the modal collapse \citep{fitting}[Sections 11.9 and 11.10]. Koons \citep{koons}, Anderson \citep{and90}[p. 292] and Sobel \citep{sobel2}[p. 133] also discuss the idea that the notion of property over which quantification is allowed might be too general and restrictions might be appropriate. Two alternative and more conservative solutions, requiring only modifications of the axioms and definitions but no change in the logic, are discussed in greater detail in the next sections.

@@

\begin{prooftree}
\AXC{$\subproof_6$} \dashedLine
\UIC{$G(\gamma)$} \RightLabel{}\dottedLine \RightLabel{D\ref{A:D1}}
\UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
\UIC{$P(\lambda x.\rho(x)) \biimp \nec \rho(\gamma)$}
\AXC{$ $} \RightLabel{4}
\UIC{$\nec\rho(\gamma)$} \RightLabel{}
\BIC{$P(\rho)$} \RightLabel{$\imp_E$}

@@

\begin{prooftree}
\AXC{$ $} \RightLabel{3}
\UIC{$\nec P(\rho)$} \RightLabel{$\nec_E$}
\UIC{$\omega: P(\rho)$}
\AXC{$ $} \RightLabel{2}
\UIC{$\omega: G(y)$}  \dottedLine \RightLabel{D\ref{A:D1}}
\UIC{$\omega: \forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
\UIC{$\omega: P(\rho) \biimp \nec \rho(y)$}\RightLabel{$\biimp_E$}
\BIC{$\omega: \nec \rho(y)$}\RightLabel{$\nec_E$}
\UIC{$\omega: \rho(y)$} \RightLabel{$\imp_I^2$}
\UIC{$\omega: G(y) \imp \rho(y)$}\RightLabel{$\forall_I$}
\UIC{$\omega: \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\nec_I$}
\UIC{$\nec \forall y.(G(y) \imp \rho(y))$} \RightLabel{$\imp_I^3$}
\UIC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$} \RightLabel{$\imp_E$}

\AXC{$\subproof_7$} \dashedLine
\UIC{$\nec P(\rho)$}
\BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^4$}
\UIC{$[\subproof_8] \qquad \nec\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
\end{prooftree}
